---
title: "Week6_Dataframes_Matrices"
author: "Harshal Shah"
date: "Dec 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Chapter 8: Matrices and Dataframes

The following table shows the results of a survey of 10 pirates. In addition to some basic demographic information, the survey asked each pirate "What is your favorite superhero?"" and "How many tattoos do you have?""

```{r echo = FALSE}
superhero <- data.frame(
  Name = c("Astrid", "Lea", "Sarina", "Remon", "Letizia", "Babice", "Jonas", "Wendy", "Niveditha", "Gioia"),
  Sex = c("F", "F", "F", "M", "F", "F", "M", "F", "F", "F"),
  Age = c(30, 25, 25, 29, 22, 22, 35, 19, 32, 21),
  Superhero = c("Batman", "Superman", "Batman", "Spiderman", "Batman",
               "Antman", "Batman", "Superman", "Maggott", "Superman"),
  Tattoos = c(11, 15, 12, 5, 65, 3, 9, 13, 900, 0)
)
knitr::kable(superhero)
## Creating the  dataframe
```


1. Combine the data into a single dataframe. Complete all the following exercises from the dataframe!
  ## We combine dataframe in to a single function by calling data.frame function. Important Property here is stringasfactor
```{r}
## This combines all data in a single frame. Make sure the stringasfactors is set to false
ps <- data.frame(
  name = c("Astrid", "Lea", "Sarina", "Remon", "Letizia", "Babice", "Jonas", "Wendy", "Niveditha", "Gioia"),
  sex = c("F", "F", "F", "M", "F", "F", "M", "F", "F", "F"),
  age = c(30, 25, 25, 29, 22, 22, 35, 19, 32, 21),
  superhero = c("Batman", "Superman", "Batman", "Spiderman", "Batman",
               "Antman", "Batman", "Superman", "Maggott", "Superman"),
  tattoos = c(11, 15, 12, 5, 65, 3, 9, 13, 900, 0),
  stringsAsFactors = FALSE
) 

```

2. What is the median age of the 10 pirates?
  ##median function is used to calculate the median
```{r}
##Finding the median age by median function
median(ps$age)
```

  
3. What was the mean age of female and male pirates separately?
## There are various ways we can do this one with mean function, with function, subset
```{r}
## Finding the mean by mean function
mean(ps$age[ps$sex == "F"])
mean(ps$age[ps$sex == "M"])
## OR
with(ps, 
     mean(age[sex == "F"]))
with(ps, 
     mean(age[sex == "M"]))
## OR
mean(subset(ps,
       subset = sex == "F")$age)
mean(subset(ps,
       subset = sex == "M")$age)
```

  
4. What was the most number of tattoos owned by a male pirate?
## We can do it by the following two ways 
```{r}
## The with function and max function is used to find the maximum value
with(ps, 
     max(tattoos[sex == "M"]))
# OR
max(subset(ps, 
           subset = sex == "M")$tattoos)
```


5. What percent of pirates under the age of 25 were female?

```{r}
## We find the mean by mean function and then adding the constraints in the function
with(ps, 
     mean(sex[age < 25] == "F"))
```

6. What percent of female pirates are under the age of 25?
  ##calculate using mean
```{r}
## We find the mean of the function and then add constrainst to find the percent value
with(ps, 
     mean(age[sex == "F"] < 25))
```

7. Add a new column to the dataframe called `tattoos.per.year` which shows how many tattoos each pirate has for each year in their life.

```{r}
## Adding a new column with function and then assigneing it to he data set
ps$tattoos.per.year <- with(ps, tattoos / age)
```

8. Which pirate had the most number of tattoos per year?

```{r}
## Finding the max by max function
ps$name[ps$tattoos.per.year == max(ps$tattoos.per.year)]
```

9. What are the names of the female pirates whose favorite piratesurvey is Superman?
  
```{r}
## Finding the name of the female pirates by with function
ps$name[with(ps, sex == "F" & superhero == "Superman")]
```
  
10. What was the median number of tattoos of pirates over the age of 20 whose favorite piratesurvey is Spiderman?

```{r}
## finding age over 20 for spiderman
with(ps, (tattoos[age > 20 & superhero == "Spiderman"]))
```